---
title: "Work in Progress Report"
author: "Sarah Morris"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## FINAL POSTER CODE: IST 719
# AUTHOR: Sarah Morris


Table of Contents:

1. Data Loading and Filtering
2. EDA and Univariate Plots
3. Multivariate Plots


#Necessary Libraries

```{r}
library(extrafont)
library(ggplot2)
library(RColorBrewer)
library(maps)
library(RColorBrewer)
library(forcats)
library(countrycode)
library(ggridges)
library(scales)
library(extrafont)
```



# Read Data

```{r}
population <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-08-22/population.csv')
population
```


```{r}
#find countries of origin with most refugees
refugees_coo = aggregate(population$refugees, list(population$coo_iso), sum)
colnames(refugees_coo) = c('iso3c', 'refugees')
refugees_coo
```


# Filtering

```{r}
#aggregate by year
refugees_years = aggregate(population$refugees, list(population$year), sum)
colnames(refugees_years) = c('year', 'refugees')
refugees_years
```


```{r}
#find countries of origin with most refugees
refugees_coo = aggregate(population$refugees, list(population$coo_iso), sum)
colnames(refugees_coo) = c('iso3c', 'refugees')
refugees_coo
```


```{r}

#plot of flows
population
pop.df = aggregate(population$refugees,
                     list(population$coo_name, population$coa_name), sum)
colnames(pop.df) = c("Country of Origin", "Country of Acceptance", "Number of Refugees")
pop.df
```



```{r}
#find countries of origin with most refugees
refugees_coo = aggregate(population$refugees, list(population$coo_iso), sum)
colnames(refugees_coo) = c('iso3c', 'refugees')
refugees_coo
```

```{r}
refugees_coo = aggregate(population$refugees, list(population$coo_name), sum)
colnames(refugees_coo) = c('region', 'refugees')
#refugees_coo
world_map <- map_data(map = "world")

worldfinal <- merge(world_map, refugees_coo, by = "region")
worldfinal
```


```{r}
#subset for idps
idps= population[population$idps > 0, ]
idps
```

```{r}

world_map <- map_data("world")

# Add a new column for the three-letter ISO code
world_map$iso3c <- countrycode(world_map$region, origin = "country.name", destination = "iso3c")
world_map
```


```{r}
#group by country
refugees = aggregate(population$refugees, list(population$coo_name), sum)
refugees = refugees[refugees$x > 1000000, ]
```


```{r}
#plot of flows
population
pop.df = aggregate(population$refugees,
                     list(population$coo_name, population$coa_name), sum)
colnames(pop.df) = c("Country of Origin", "Country of Acceptance", "Number of Refugees")
pop.df
```




# MERGE DATASET WITH REGIONS/SUBREGIONS

```{r}
refugees_coa = aggregate(population$refugees, list(population$coa_iso), sum)
colnames(refugees_coo) = c('iso3c', 'refugees')
#refugees_coo
world_map <- map_data(map = "world")
world_map$iso3c <- countrycode(world_map$region, origin = "country.name", destination = "iso3c")
refugees_coa
world_map
```

```{r}

all <- read.csv("~/Downloads/all.csv")

all$iso3c = all$alpha.3

#merge with population
all = all[,c("iso3c", "sub.region", "region")]
all
```

```{r}
# country code
population$iso3c = population$coo_iso
regions = merge(all, population, by = "iso3c")
regions
```


```{r}
#by region
region.df = aggregate(regions$refugees,
                     list(regions$coo_name, regions$coa_name, regions$sub.region), sum)
colnames(region.df) = c("Country of Origin", "Country of Acceptance", "Country of Origin Region", "Refugees")
region.df
```



```{r}
# filter where num refugees is over 3,000,000
region.sig = subset(region.df, Refugees >3000000)
region.sig
```



# EDA/UNIVARIATE PLOTS 

```{r}
par(family = "Times New Roman")
ggplot(refugees_years) + aes(y = refugees, x = year) + geom_line(color = "lightblue") + geom_point(color = "darkblue") + scale_x_continuous(breaks = seq(2010,2022,1)) + scale_y_continuous(breaks=c(0, 5000000, 10000000,15000000,20000000,25000000,30000000),
                     labels=c("0", "5 Million", "10 Million", "15 Million", "20 Million", "25 Million", "30 Million")) + labs(title = "Number of Refugees Worldwide", x = "Year", y = "Number of Refugees") + theme_bw() + theme(text=element_text(size=14,  family="serif"))
```


```{r}
#view distribution
hist(refugees_coo$refugees)
#data ranges between 0 and 60 mil
sort(refugees_coo$refugees)
boxplot(refugees_coo$refugees)
```

# WORD CLOUD

```{r}
library(wordcloud2)
library(dplyr)

word_freq <- population %>%
  group_by(word = coo_name) %>%
  summarise(freq = n()) %>%
  ungroup()%>%  arrange(desc(freq))

blue_gradient <- c("#003b64", "#00639b", "#0087ca", "#00ddf9", "#00abe7")

wordcloud2(data = word_freq, 
            size = 1,
            color = blue_gradient,
            backgroundColor = "#2e2e2e")
```


# IDPS DENSITY PLOT 

```{r}
ggplot(idps, aes(x = idps, y = 0, fill = after_stat(x))) + 
  geom_density_ridges_gradient()+
  scale_fill_gradient2(low="lightblue1", mid = "lightblue2", high="blue") +
  scale_x_continuous(labels = comma) +  # Format x-axis labels with commas
  scale_y_continuous(labels = percent) +
  labs(title = "Density Plot of Internally Displaced Persons by Country", subtitle = "*Includes only Countries with >0 internally displaced persons" , x = "Total IDPs", y = "Countries", fill = "IDPs") + theme(
    text = element_text(family = "Times New Roman"),
    panel.background = element_rect(fill = "#2E2E2E"),  # This can be set to black if desired
    plot.background = element_rect(fill = "#2E2E2E"),
    panel.grid.major = element_blank(),  # Hide major grid lines
    panel.grid.minor = element_blank(),
    plot.title = element_text(color = "white"),        # Title text color
    plot.subtitle = element_text(color = "lightgrey"),
    axis.title = element_text(color = "white"),        # Axis titles color
    axis.text = element_text(color = "white"))  # Hide minor grid lines
```


# REFUGEE DENSITY PLOT

```{r}
#install.packages("ggridges")
#install.packages("scales")
library(ggridges)
library(scales)
library(extrafont)

ggplot(refugees, aes(x = x, y = 0, fill = after_stat(x))) + 
  geom_density_ridges_gradient()+
  scale_fill_gradient2(low="lightblue1", mid = "lightblue2", high="blue") +
  scale_x_continuous(labels = comma) +  # Format x-axis labels with commas
  scale_y_continuous(labels = percent) +
  labs(title = "Density Plot of Refugee Counts by Country", subtitle = "*Includes only Countries with >1,000,000 Refugees" , x = "Total Refugees", y = "Countries", fill = "Refugees") + theme(
    text = element_text(family = "Times New Roman"),
    panel.background = element_rect(fill = "#2E2E2E"),  # This can be set to black if desired
    plot.background = element_rect(fill = "#2E2E2E"),
    panel.grid.major = element_blank(),  # Hide major grid lines
    panel.grid.minor = element_blank(),
    plot.title = element_text(color = "white"),        # Title text color
    plot.subtitle = element_text(color = "lightgrey"),
    axis.title = element_text(color = "white"),        # Axis titles color
    axis.text = element_text(color = "white"))  # Hide minor grid lines
  
```




## MULTIVARIATE PLOTS


# GGPLOT MAP - country of origin

```{r}
plain <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(hjust = 0.5)
)

worldfinal$refugee_group <- cut(worldfinal$refugees, 
                      breaks = c(-Inf, 1000000, 5000000, 10000000, 15000000, 20000000, 50000000, Inf), 
                      labels = c("Less than 1 Million", "1 Million to 5 Million", "5 Million to 10 Million", "10 Million to 15 Million", "15 Million to 20 Million", "20 Million to 50 Million", "More than 50 Million"))

ggplot(data = worldfinal, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = NA, color = "black") +
  geom_polygon(aes(fill = refugee_group)) +
  scale_fill_manual(name = "Counts", values = rev(brewer.pal(7, name = "Blues"))) + 
  ggtitle("Refugee Countries of Origin") + plain
```



## GGPLOT COUNTRIES OF ACCEPTANCE

```{r}
refugees_coa = aggregate(population$refugees, list(population$coa_iso), sum)
colnames(refugees_coa) = c('iso3c', 'refugees')
#refugees_coo
world_map <- map_data(map = "world")
world_map$iso3c <- countrycode(world_map$region, origin = "country.name", destination = "iso3c")

worldfinal2 <- merge(world_map, refugees_coa, by = "iso3c")
worldfinal2



plain <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(hjust = 0.5)
)

worldfinal2$refugee_group <- cut(worldfinal2$refugees, 
                      breaks = c(-Inf, 1000000, 5000000, 10000000, 15000000, 20000000, 50000000, Inf), 
                      labels = c("Less than 1 Million", "1 Million to 5 Million", "5 Million to 10 Million", "10 Million to 15 Million", "15 Million to 20 Million", "20 Million to 50 Million", "More than 50 Million"))

ggplot(data = worldfinal2, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = NA, color = "black") +
  geom_polygon(aes(fill = refugee_group)) +
  scale_fill_manual(name = "Counts", values = rev(brewer.pal(7, name = "Blues"))) + 
  ggtitle("Refugee Countries of Acceptance") + plain
```




## ALLUVIAL PLOT



# ALluvial plot of largest flows of refugees

```{r}
par(family = "Times New Roman")

#alluvial plot with country oo to country of oa 
library(alluvial)

# Define colors for each country in "Country of Acceptance"
num_regions = length(unique(region.sig$`Country of Origin Region`)) #9 unique
colors <- brewer.pal(min(num_regions, 9), "Blues")
region_names <- unique(region.sig$`Country of Origin Region`)
region_colors <- setNames(colors, region_names)




alluvial(
  region.sig[, 1:2],
  freq = region.sig$Refugees,
  col = region_colors[region.sig$`Country of Origin Region`], 
  cex = 0.7
  
)

```

# Create Legend Isolated

```{r}
# Create a blank alluvial plot to isolate legend
alluvial(
  region.sig[, 1:2],
  freq = region.sig$Refugees,
  col = region_colors["white"], 
  cex = 0.7
)

legend(
  x = 1.5,  
  y = 0.8,  
  legend = names(region_colors),  
  fill = region_colors,  
  title = "Country of Origin Region",  
  cex = 0.8,  
  bty = "y",  
  bg = "#2E2E2E",
  text.col = "white"
)
```



population

```{r}
word_freq <- population %>%
  group_by(word = coo_name) %>%
  summarise(freq = n()) %>%
  ungroup()%>%  arrange(desc(freq))
#word_freq

refugees%>% arrange(desc(x))

Syria - almost 60M
Afghanistan - 40M
South Sudan - 17M
Somalia - 13M
Myanmar - 10M



```
```{r}
origin = aggregate(population$refugees, list(population$coa_name), sum)

origin%>% arrange(desc(x))

Turkey - 30M
Pakistan - 20M
Iran - 15M
Germany - 10M
Uganda - 10M

```

